Smart Representations: {R}ationality and Evolution in a Richer Environment

### 2015-12-18 paper/paper.tex
rewrite sec com

\emph{Meta games} are abstract models for the evolutionary competition between choice
mechanisms in social decision making contexts. The expected payoff for a $c$-player in a random
encounter with a $c'$-player is the average of payoffs obtained in a random game $g$ weighted
by $g$'s occurrence probability:
\begin{align}
  F(c, c') = \int P(g) \  \utils_g(c,c') \text{ d} g \,.
\end{align}

@@

Standard notions of evolutionary game theory can be applied to meta games as well. A choice
principle $c$ is a strict Nash equilibrium if $F(c,c) > F(c',c)$ for all $c'$; it is
evolutionarily stable if for all $c'$ either (i) $F(c,c) > F(c',c)$ or (ii) $F(c,c) = F(c',c)$
and $F(c,c') > F(c',c')$. It is neutrally stable if for all $c'$ either (i) $F(c,c) > F(c',c)$
or (ii) $F(c,c) = F(c',c)$ and $F(c,c') \ge F(c',c')$. Similarly, evolutionary dynamics can be
applied to meta games. We will later turn towards the replicator dynamic
\citep{TaylorJonker1978:Evolutionary-St} and the replicator mutator dynamic
\citep[e.g.][]{Nowak2006:Evolutionary-Dy}.\todo{footnote with caveat about agent-level
  processes}

@@

Maximinimization under a precise (singleton set) probabilistic belief is just standard
maximization of expected utilities. Moreover, given the same precise (singleton set)
probabilistic belief, the choices of objective and regret preference types are exactly the
same.

\begin{fact} \label{fact:maxEU-minReg} For any strategic game $G$ and any arbitrary precise
  (singleton set) probabilistic belief about the opponents' behavior, the acts selected by
  maximization of expected utility for $G$ are exactly the acts selected by the maximization of
  expected utility for the negative regret transformation of $G$
  \citep[e.g.,][]{HalpernPass2012:Iterated-Regret}. This also holds for solitary decision
  problems.
\end{fact}

This means that as far as behavior is concerned, regret types cannot be distinguished from
objective types, as long as both hold precise beliefs. This will be central in understanding
later results: under precise beliefs evolution will not favor either preference type over the
other. On the other hand, if paired with imprecise beliefs, different behavioral predictions
arise. However, in the special case of $2 \times 2$ symmetric games, another noteworthy general
result holds.

@@

In the following, we assume that agents are maximally ignorant, so that precise types have a
(singleton set containing a) flat probability distribution
$\overline{\mu}_i \in \Delta(\Acts_{-i})$ and imprecise types consider all probabilistic
beliefs about the co-players' behavior $\overline{X} = \Delta(\Acts_{-i})$. In that case, the
considered choice mechanisms are behaviorally equivalent to classic choice rules, as summarized
in Table~\ref{tab:CMs}. This means that most of the following results which we would like to
interpret as results about the evolutionary fitness of subjective representation schemes have
value also for anyone interested only in a comparison of the ecological value of classic choice
rules. This also means that, given Fact~\ref{fact:equivalence2x2}, all choice mechanisms are
behaviorally equivalent on $2 \times 2$ symmetric games, except for objective preference types
with imprecise beliefs.

\subsection{Imprecise probabilistic beliefs}
\label{sec:impr-prob-beli}

Consider the following problem. I have, on the desk in front of me, a bag containing either red or black marbles. There is no further information about the distribution of the two colors. What is then the probability that I will draw a red marble? As in \citet{walley96}, we are not talking about physical probability, but rather about epistemic probability, i.e., the subjective probability that an agent attributes to an event.
In the literature there are two traditional ways of answering this kind of questions. According to standard Bayesianism, one should always be able to determine a precise betting rate (that immediately translates into a precise probability distribution), which is fair, in the sense that the DM is indifferent between betting on or against red at the fair rate, and would bet on red at any rate that is more favourable than the fair rate and against red at any rate that is less favourable. In the case of the bag on my desk, a strict Bayesian, in the absence of any further information, would probably rely on the \textit{principle of insufficient reason} and answer that the DM should have a uniform distribution that considers all the outcomes equipossible. \\
The second answer involves imprecise probabilities and is related to the literature about unmeasurable uncertainty. Many authors argued in favor of this kind of approach \citep[e.g.,][]{levi74,gardsah82,walley96}. Since it is consistent with the information available that all the marbles are red, or that no marble is red, the resulting model should take it into account and specify a lower probability of red $\underline{P}(R)$ and an upper probability of red $\overline{P}(R)$ that define an interval of possible probabilistic beliefs [$\underline{P}(R), \overline{P}(R)$]. Given the information available in this case we would have $\underline{P}(R)=0$ and $\overline{P}(R)=1$. This approach appears particularly relevant in a game theoretical context. Indeed, in a recent paper \citet{BattCerrMM15} write:

\begin{quote}
Such uncertainty is inherent in situations of strategic interaction. This is quite obvious when such situations have been faced only a few times.
\end{quote}

In evolutionary game theory, for instance, players in a population obviously face uncertainty about the composition of the population that they are part of, and consequently about the co-player that they are randomly paired with at each round and about the co-players' action. We represent this uncertainty in two different ways that correspond to the two approaches discussed above: a Bayesian uniform distribution and an imprecise probability set. This amounts to having two different belief types, $ \overline{\mu}$ and $ \Delta $, where $\overline{\mu}$ stands for the uniform distribution and $\Delta$ is the probability set, defined as follows:

\item $ \Delta \equiv \Delta(A_{-i}) :=\lbrace \mu \in \mathbb{R}^{A_{-i}}: \sum_{a_{-i} \in A_{-i}} \mu(a_{-i}) = 1 \rbrace$;

\item $\overline{\mu} \in \Delta$ is the probability measure over the co-players' actions such that: $\forall a_{-i} \in A_{-i}$, $\overline{\mu}(a_{-i})= \frac{1}{|A_{-i}|}$.

Hence, we consider two possible epistemic states in the face of uncertainty. The belief type $\Delta$ represents a situation of non-Bayesian radical uncertainty, where the players are not able to narrow down the set of probability distributions to a single one and they consider all of them as possible. The other belief type $\overline{\mu}$ conceptualizes the uncertainty about the population by using the principle of insufficient reason and ascribes equal probability to all the possible alternatives. As we have seen, both options are reasonable and justifiable ways to deal with situations of uncertainty. \\

In decision theory, the imprecise probability model is in line with most of the representation results of decision making under uncertainty \citep[e.g.,][]{gilsch89,KlibMarMuk05,GhirMar02}, and seems justified by empirical observations too.

A prominent example is Ellsberg's paradox (\citet{ells61}). A bag contains 90 marbles, 30 are red and the remaining 60 are either black or yellow. A marble is to be drawn and DM can win 100€ if she guesses the color of the drawn marble. The DM is offered two different bets. In the first one she can choose to bet either on red (R) or on yellow (Y), while in the second she can bet either on red or black (RB) or on yellow or black (YB), as shown in Table \ref{Ellsberg}.

The pattern consistently observed in experimets is: $f_{{\color{red}R}} \succ f_{{\color{yellow}Y}}$ and $f_{{\color{red}R}B} \prec f_{{\color{yellow}Y}B}$. This is clearly incompatible with any precise probabilistic belief about the proportion of black and yellow marbles. This kind of behavior has been
famously axiomatized by Gilboa and Schmeidler in a paper from 1989. The choice pattern is normally explained in terms of aversion to uncertainty, and the standard representation is given by means of a set of probability measures together with maxmin rule. We can think of the DM in Ellsberg's example as maximinimizing expected utility over the probability set $ [\underline{P}(R)=\overline{P}(R)=\frac{1}{3},\underline{P}(Y)=0,\underline{P}(B)=0,\overline{P}(Y)=\frac{2}{3},\overline{P}(B)=\frac{2}{3}]  $. This amounts to calculating for any probability measure contained in the probability set the expected utility of any available action, and then to choosing the action that guarantees the highest minimal expected utility. The following observation is then trivial.

\begin{fact} \label{fact:singleton probability set}
Whenever the probability set is a singleton, maximinimization of
expected utility coincides with maximization of expected utility.
\end{fact}

\section{Meta-games and the evolution of choice mechanisms}
\label{sec:model}

The standard approach in
evolutionary game theory is not enough for our purposes, because it (i) fixes a unique and immutable
\emph{stage game} (e.g., the prisoner's dilemma) and (ii) assumes that the object of
evolutionary selection is behavior for that very game. Following a line of argument coming from theoretical biology and behavioral ecology
\citep[e.g.,][]{FawcettHamblin2013:Exposing-the-be}, we therefore turn towards the evolutionary
competition of \emph{choice mechanisms}, i.e., general ways of selecting concrete behavior
across a range of different situations, defined by the two functions $\gamma$ and $\kappa$. With this approach, the expected fitness of a choice
mechanism will not be determined by a unique stage game, but it will be a
complex function of the statistical properties of the environment: we have to take into account
how frequently particular concrete situations of choice making are.

For concreteness, let us fix a class of strategic games $\mathcal{G}$ and a probability measure
$P \in \Delta(\mathcal{G})$ that gives the occurrence probability $P(G)$ of each game
$G \in \mathcal{G}$. Together, a class of games and a probability measure over it can be
thought of as a model of the environment in which different choice mechanisms compete. As we have seen in Definition \ref{def:maxminSEU}, given a game $G \in \mathcal{G}$, a choice mechanism prescribes the action(s) $\kappa(\gamma(G))$. Now imagine a huge, possibly infinite society of agents, each of which carries a choice mechanism, and a class $\mathcal{G}$ of symmetric two player games for simplicity.
Definition \ref{def:expected evolutionary payoff} gives the expected evolutionary payoff of choice mechanism $\kappa(\tau, e)$ against choice mechanism $\kappa(\tau', e')$ in game $G \in \mathcal{G}$.

\begin{definition} \label{def:expected evolutionary payoff}

Given a game $G \in \mathcal{G}$, a choice mechanism $\kappa(\tau, e)$, and a choice mechanism $\kappa(\tau', e')$, we define $$\pi^G(\kappa(\tau(G), e),\kappa(\tau'(G), e')):= \sum_{a_i \in \kappa(\tau(G), e)} \sum_{a_j \in \kappa(\tau'(G), e')} \pi^G(a_i , a_j) \cdot \frac{1}{|\kappa(\tau(G), e)|} \cdot \frac{1}{|\kappa(\tau'(G), e')|}.$$

\noindent Then the expected evolutionary fitness $EF_{\mathcal{G}}(\kappa(\tau, e))$ of an agent with
choice mechanism $\kappa(\tau, e)$ is determined by the probability $P((\tau',e'))$ of meeting an agent with
choice mechanism $\kappa(\tau',e')$, which we will assume is the frequency of $(\tau',e')$ in the population, the
probability $P(G)$ of playing game $G \in \mathcal{G}$ and the evolutionary payoff
$\pi^G(\kappa(\tau, e),\kappa(\tau', e'))$:

\begin{align}
\label{eq:EF}
  EF_{\mathcal{G}}(\kappa(\tau, e)) := \sum_{G \in \mathcal{G}} \sum_{(\tau',e') \in T \times E} P(G) \cdot P((\tau',e')) \cdot \pi^G (\kappa(\tau(G), e),\kappa(\tau'(G), e'))\,.
\end{align}

\noindent Similarly, the expected fitness of a choice mechanism $\kappa(\tau, e)$, when playing against another choice mechanism $\kappa(\tau', e')$ is:
\begin{align}
  \label{eq:pairwiseEF}
  EF_{\mathcal{G}}(\kappa(\tau, e),\kappa(\tau', e')) := \sum_{G \in \mathcal{G}}  P(G) \cdot \pi^G(\kappa(\tau(G), e),\kappa(\tau'(G), e'))\,.
\end{align}

Notice that the expected fitness of a choice mechanism is defined entirely in parallel to the
expected payoff of an act in a fixed stage game (Definition \ref{def:expected evolutionary payoff}), except that where the latter has a simple
payoff function, the former has an expectation over payoffs in a statistically rich
environment. That means that we can think of evolutionary competition between choice mechanisms
in terms of payoffs defined by a \emph{meta-game} over $\mathcal{G}$ whose ``acts'' are choice
mechanisms and whose ``payoffs'' are the expected evolutionary payoffs of
Equation~(\ref{eq:pairwiseEF}).\footnote{Allowing games in $\mathcal{G}$ with different numbers of players makes matters formally, but not conceptually, more complicated. In the following we will focus entirely on two player games, hence we can neglect further complications for the moment.}

@@

To demonstrate the usefulness of a meta-game approach, let us concentrate first on a basic
model that contains the four choice mechanisms obtained from crossing the preference and belief
types introduced in Section~\ref{sec:basic-notions}.  For particular empirical purposes, one
could consult a more specific class $\mathcal{G}$ with appropriate, possibly empirically
determined probabilities $P(G)$ in order to match the natural environment of a given
population. But for our theoretical purposes we adopt another approach.

For a first basic
model, let $\mathcal{G}$ be a set of symmetric two player games with two acts. (We will look at
extensions and generalizations in Section~\ref{sec:extensions}). More concretely, we imagine the
class $\mathcal{G}$ to be generated from the generic form table of Table \ref{generic form
  table}, where $a,b,c,d$ are i.i.d.~random variables sampled from a finite set of natural
numbers $ \lbrace 0, \dots, 10 \rbrace$. After a game $G \in \mathcal{G}$ is generated, players in the population are randomly paired to play the game $G$. Each player represents a choice mechanism, and selects an action accordingly. At the end of the game, the evolutionary payoffs of each player are recorded, another game $G' \in \mathcal{G}$ is generated, players are randomly paired again, and so on.

\iffalse
\begin{figure}
\begin{center}%
\begin{tabular}{|c|c|c|}
\hline 
 & I & II\tabularnewline
\hline 
\hline 
I & a;a & b;c\tabularnewline
\hline 
II & c;b & d;d\tabularnewline
\hline 
\end{tabular}\end{center}

\protect\caption{generic form table for $2 \times 2$ symmetric games}
\label{generic form table}
\end{figure}
\fi

\begin{table}[]
\centering
%\caption{generic form table for $2 \times 2$ symmetric games}
%\label{generic form table}
\begin{tabular}{|l|l|l|}
\hline
 & $I$ & $II$ \\ \hline
$I$ & $a;a$ & $b;c$ \\ \hline
$II$ & $c;b$ & $d;d$ \\ \hline
\end{tabular}
\caption{Generic form table for $2 \times 2$ symmetric games}
\label{generic form table}
\end{table}

\subsection{Preference types}

The set of all possible preference types $T$ amounts to the set of all possible functions $ \tau: \mathcal{G} \rightarrow  \Theta $, i.e., $\tau: \mathbb{R}^{4} \rightarrow \mathbb{R}^{4} $. However, here we want to focus on some subsets of $T$ that are also meaningful from a decision-theoretic point of view. As introduced in Section \ref{sec:basic-notions}, we are mainly interested in two types: the objective type and the regret type. In other words, we consider two prominent preference types: a type who has a veridical subjective utility that perfectly matches the objective utility (i.e., the evolutionary fitness), and a type who has a non-veridical regret-based subjective utility.

Given a game $G$ and Definition \ref{defn:parwreg}, it is straightforward to define the
\textit{regret preference} $\theta^{reg}$ and the \textit{regret type} $\tau^{reg}$ as
follows.

\begin{definition}[Regret preference] \label{defn:regpref}

For a game $ G=\langle N, (A_i , \pi_i)_{i \in N} \rangle $, we define the regret preference $ \theta^{reg}: A \rightarrow \mathbb{R} $ such that $ \theta^{reg}(a_i,a_{-i})=-\text{reg}(a_i,a_{-i}) $.\footnote{More precisely, this is the negative regret, whereas the (positive) regret, as defined in [halpass], would not have the minus sign at the beginning of the formula. Here it will be easier to define the regret in terms of negative regret.}

\begin{definition}[Regret type] \label{defn:regtype}

Given $ \mathcal{G} $ the class of symmetric $ 2 \times 2 $ games, we call regret type the function $\tau^{reg}: \mathcal{G} \rightarrow  \Theta$ such that, for all $G \in \mathcal{G}$, $ \tau^{reg}(G)= \theta^{reg}$.

\noindent Similarly, let us define the \textit{objective preference} $\theta^{\pi}$ and the \textit{objective type} $ \tau^{\pi} $, that we also call $\pi$\textit{-type}, as follows.

\begin{definition}[Objective preference] \label{defn:objpref}

For a game $ G=\langle N, (A_i , \pi_i)_{i \in N} \rangle $, we define the objective preference $ \theta^{\pi}: A \rightarrow \mathbb{R} $ such that $ \theta^{\pi}(a_i,a_{-i})=\pi(a_i,a_{-i}) $.

\begin{definition}[Objective type] \label{defn:objtype}

Given $ \mathcal{G} $ the class of symmetric $ 2 \times 2 $ games, we call objective type, or $\pi$-type, the function $\tau^{\pi}: \mathcal{G} \rightarrow  \Theta$ such that, for all $G \in \mathcal{G}$, $ \tau^{\pi}(G)= \theta^{\pi}$.

\section{Results for the basic model}
\label{sec:results:-basic-model}

In this section we present results of computer simulations and analytical proofs for the basic
model with the four choice mechanisms introduced in the previous sections.

\subsection{Numerical simulations}
\label{sec:numer-simul}

Table~\ref{tab:ExpectedFitness_4Types} gives an approximation of expected fitness, obtained by
numerical simulation from averaging over $100,000$ randomly sampled games in $\mathcal{G}$ with
$a,b,c,d \in \lbrace 0, \dots, 10 \rbrace$. Games were sampled repeatedly by
choosing independently four integers $a,b,c,d \in \lbrace 0, \dots, 10 \rbrace$ uniformly at
random. For each game, a player type's action choices were determined and payoffs from all
pairwise encounters recorded. (Whenever a player type would not select a unique action in a
given game, we recorded unbiased averages over playing either action.) The number in each cell
of Table~\ref{tab:ExpectedFitness_4Types} is the average payoff for the row player type over
all 100,000 numbers for each pairing of player types obtained in this way. These averages
approximate the corresponding average evolutionary fitness of the row player when matched with
the column player.

\begin{table}[t]
\centering
\begin{tabular}{ccccc}
  \toprule
 & $\tau^{reg}, \Delta$ 
 & $\tau^{\pi}, \Delta$ 
 & $\tau^{reg}, \overline{\mu}$ 
 & $\tau^{\pi}, \overline{\mu}$ \\ 
  \midrule
  $\tau^{reg}, \Delta$ & 6.663 & 6.662 & 6.663 & 6.663 \\ 
  $\tau^{\pi}, \Delta$ & 6.486 & 6.484 & 6.486 & 6.486 \\ 
  $\tau^{reg}, \overline{\mu}$ & 6.663 & 6.662 & 6.663 & 6.663 \\  
  $\tau^{\pi}, \overline{\mu}$ & 6.663 & 6.662 & 6.663 & 6.663 \\ 
   \bottomrule
\end{tabular}                    
\caption{Average evolutionary fitness in simulations of 100,000 symmetric $2 \times 2$ games}
\label{tab:ExpectedFitness_4Types}
\end{table}

Simulation results obviously reflect Fact~\ref{fact:equivalence2x2} in that all encounters in
which types $(\tau^{reg}, \Delta)$, $(\tau^{reg}, \overline{\mu})$ or
$(\tau^{\pi}, \overline{\mu})$ are substituted for one another yield identical results. More
interestingly, Table~\ref{tab:ExpectedFitness_4Types} shows that the maxmin strategy
$(\tau^{\pi}, \Delta)$ is strictly dominated by either of the three other strategies:
in each column (i.e., for each kind of opponent), the maxmin strategy is strictly worse
than any of the three behaviorally equivalent competitors. This has a number of interesting
consequences.

Let's begin with a restricted scenario and look at more and more complex cases afterwards. If
we restrict attention to a population of radically uncertain players, i.e., where only
belief type $\Delta$ exists, a population of regret types $ \tau^{reg} $ is the only
\emph{evolutionarily stable state}. An evolutionarily stable state is a state which cannot be
invaded by a small number of mutants. More strongly, since maxmin play is strictly dominated,
we expect selection that is driven by expected fitness to invariably weed out maxmin behavior
in favor of regret minimization. In other words, when taking average payoffs over the class of
random games considered here, regret minimization is a better choice mechanism than standard maxmin
from an evolutionary point of view, despite the fact that the former, but not the latter,
entertains non-veridical subjective preferences.

Next, if we look at the competition between all four types represented in
Table~\ref{tab:ExpectedFitness_4Types}, $(\tau^{reg}, \Delta)$ is no longer
evolutionarily stable. Rather, given behavioral equivalence (Fact~\ref{fact:equivalence2x2}),
types $(\tau^{reg}, \Delta)$, $(\tau^{reg}, \overline{\mu})$, and
$(\tau^{\pi}, \overline{\mu})$ are all \emph{neutrally stable}
\citep{Maynard-Smith1982:Evolution-and-t}. But since $(\tau^{\pi}, \Delta)$ is strictly
dominated and so disfavored by fitness-based selection, we are still drawn to conclude that
maxmin behavior is weeded out in favor of a population with a random distribution of the
remaining three types.

Simulation results of the (discrete time) \emph{replicator dynamics}
\citep{TaylorJonker1978:Evolutionary-St} indeed show that random initial population
configurations are attracted to states with only three player types:
$(\tau^{reg}, \Delta)$, $(\tau^{reg}, \overline{\mu})$ and
$(\tau^{\pi}, \overline{\mu})$. The relative proportions of these depend on the initial
population.\todo{check uniform notation for player types: round vs.~square brackets}  Nonetheless, this
variability disappears if we add a small mutation rate to the dynamics. We assume a fixed,
small mutation rate $\epsilon$ for the probability that a player's preference type \emph{or}
her belief type changes to another preference type or belief type. The probability
that a player randomly mutates into a completely different player with altogether
different preference type and belief type would then be $\epsilon^2$. With these assumptions
about ``component-wise mutations'', numerical simulations of the (discrete time)
\emph{replicator mutator dynamics} \citep{Nowak2006:Evolutionary-Dy} show that already for very
small mutation rates almost all initial populations converge to a single fixed point in which
the majority of players are regret types. For instance, with $\epsilon = 0.001$, almost all
initial populations are attracted to a final distribution with proportions:\todo{recalculate
  these numbers for the actual 4x4 game!}

\begin{center}
  \begin{tabular}{ccc}
    $(\tau^{reg}, \Delta)$ & $(\tau^{reg},
      \overline{\mu})$ & $(\tau^{\pi}, \overline{\mu})$ \\ \hline
    0.279  &   0.383 &    0.281 
  \end{tabular}
\end{center}

What this suggests is that, since biological evolution selects behavior-generating mechanisms,
not behavior as such, it need not be the case that behaviorally equivalent mechanisms are
treated equally all the while. If mutation probabilities are a function of individual
components (i.e., preference and belief types) of such behavior-generating mechanisms, it can be the case that certain components
are more strongly favored by a process of random mutation and selection. This is exactly the
case with regret-based subjective preferences in the present example. Since regret-based
subjective preferences are much better in connection with radical uncertainty than veridical
preferences are, the proportion of expected regret minimizers in the attracting state is
substantially higher than that of expected utility maximizers, even though these types are
behaviorally equivalent. In sum, component-wise mutation could have lead to a majority of
regret-based players even if, in the context of two-by-two symmetric games, they are behaviorally indiscernible from decision makers with
veridical preferences, because regret types are better off under radical uncertainty.

@@

Results based on the single meta-game in Table~\ref{tab:ExpectedFitness_4Types}, which was
obtained by averaging over numerical simulations, are not fully general and possibly spoiled by
random fluctuations in the sampling procedure. Fortunately, for the case of two-by-two symmetric games, the main result that maxmin types $(\tau^{\pi}, \Delta)$ are
strictly dominated can also be shown analytically for generous general conditions on how likely
randomly sampled games are. Let $\Lambda$ denote the set of player types in the population, and fix $\kappa$ to be maxmin expected utility as usual.

\begin{proposition} \label{proposition1}

 Fix $\Lambda = \lbrace (\tau^{\pi}, \Delta), (\tau^{reg}, \Delta) \rbrace$, and let $\mathcal{G}$ be the class
  of $2 \times 2$ symmetric games with payoffs sampled from a
  set of i.i.d.~values with at least three elements in the support. Then, $(\tau^{reg}, \Delta)$ is the only (strict) Nash
  equilibrium in the resulting meta-game.

\begin{proof}
See Appendix.
\end{proof}

\noindent The next corollary about evolutionary stability follows from proposition \ref{proposition1}.

\begin{corollary} \label{corollary1}

Fix $\Lambda = \lbrace (\tau^{\pi}, \Delta), (\tau^{reg}, \Delta) \rbrace$, and let $\mathcal{G}$ be the class
  of $2 \times 2$ symmetric games with payoffs sampled from a
  set of i.i.d.~values with at least three elements in the support. Then, the unique evolutionarily stable state is a monomorphic population of $(\tau^{reg}, \Delta)$-players.

>>>>

\emph{Meta games} are abstract models for the evolutionary competition between choice
mechanisms in social decision making contexts. The expected payoff for a $c$-player in a random
encounter with a $c'$-player is the average of payoffs obtained in a random game $g$ weighted
by $g$'s occurrence probability:
\begin{align}
  \label{eq:FittnessChoiceMechGamePairwise}
  F(c, c') = \int P(g) \  \utils_g(c,c') \text{ d} g \,.
\end{align}

@@

Standard notions of evolutionary game theory can be applied to meta games as well. A choice
principle $c$ is a strict Nash equilibrium if $F(c,c) > F(c',c)$ for all $c'$; it is
evolutionarily stable if for all $c'$ either (i) $F(c,c) > F(c',c)$ or (ii) $F(c,c) = F(c',c)$
and $F(c,c') > F(c',c')$. It is neutrally stable if for all $c'$ either (i) $F(c,c) > F(c',c)$
or (ii) $F(c,c) = F(c',c)$ and $F(c,c') \ge F(c',c')$
\citep{Maynard-Smith1982:Evolution-and-t}. Similarly, evolutionary dynamics can be applied to
meta games. We will later turn towards the replicator dynamic
\citep{TaylorJonker1978:Evolutionary-St} and the replicator mutator dynamic
\citep[e.g.][]{Nowak2006:Evolutionary-Dy}.\todo{footnote with caveat about agent-level
  processes}

@@

\begin{fact} \label{fact:singleton probability set}
Maximinimization under a precise (singleton set) belief is just standard
maximization of expected utilities.
\end{fact}

\begin{fact} \label{fact:maxEU-minReg} Under an identical precise (singleton set) belief, the
  acts selected by maximization of expected utility in game $G$ are exactly the acts selected
  by the maximization of expected utility for the negative regret transformation of $G$
  \citep[e.g.,][]{HalpernPass2012:Iterated-Regret}.
\end{fact}

This means that as far as behavior is concerned, regret types cannot be distinguished from
objective types, as long as both hold precise beliefs. This will be central in understanding
later results: under precise beliefs evolution will not favor either preference type over the
other. On the other hand, if paired with imprecise beliefs, different behavioral predictions
arise.

In the following, we assume that agents are maximally uncertain, so that precise types have a
(singleton set containing a) flat probability distribution
$\overline{\mu}_i \in \Delta(\Acts_{-i})$ and imprecise types consider all probabilistic
beliefs about the co-players' behavior $\overline{X} = \Delta(\Acts_{-i})$. In that case, the
considered choice mechanisms are behaviorally equivalent to classic choice rules, as summarized
in Table~\ref{tab:CMs}. This means that most of the following results which we would like to
interpret as results about the evolutionary fitness of subjective representation schemes have
value also for anyone interested only in a comparison of the ecological value of classic choice
rules.

Maximal uncertainty also means that, given Fact~\ref{fact:equivalence2x2}, all choice
mechanisms are behaviorally equivalent on $2 \times 2$ symmetric games, except for objective
preference types with imprecise beliefs.

@@

@@

\subsection{Simulation results}
\label{sec:simulation-results}

To demonstrate the usefulness of a meta-game approach, let us concentrate first on a basic
model that contains the four choice mechanisms obtained from crossing the preference and belief
types introduced in Section~\ref{sec:basic-notions}. From now on we speak of types when we mean
pairs like $(\text{reg}, \text{imp})$, a choice mechanism with regret-based preferences and
imprecise beliefs, or $(\text{obj}, \text{prc})$, one with objective preferences and precise
beliefs.

Meta games factor in statistical properties of the environment. For a particular empirical
purposes, one would consult a specific class of games $\mathcal{G}$ with appropriate, possibly
empirically informed probabilities $P(G)$ in order to match the natural environment of a given
population. But for our theoretical purposes we adopt another approach, since we would not know
how to justify any more specific assumption. To begin with, let $\mathcal{G}$ be a set of
symmetric 2-player games with two acts. A game is then individuated solely by its payoff
function, which is essentially a matrix with four numbers. As for occurrence probabilities of
games, we imagine that entries in a game's payoff matrix are i.i.d.~random variables sampled
uniformly from set $ \lbrace 0, \dots, 10 \rbrace$.  Using Monte Carlo simulation, we can then
approximate the relevant fitness values of Equation~(\ref{eq:FittnessChoiceMechGamePairwise})
to construct meta game payoffs. Results based on $100,000$ randomly sampled games are given in
Table~\ref{tab:ExpectedFitness_4Types}.\footnote{Concretely, $100,000$ games were sampled
  repeatedly by choosing independently four integers between 0 and 10 uniformly at random. For
  each game, the action choices of all four choice mechanisms were determined and payoffs from
  all pairwise encounters recorded. (Whenever a choice mechanism would not select a unique
  action in a given game, we recorded unbiased averages over all selected actions.)  The number
  in each cell of Table~\ref{tab:ExpectedFitness_4Types} is the average payoff for the choice
  mechanism listed in the row when matched with the choice mechanism in the column.}

\begin{table}[t]
\centering
\begin{tabular}{ccccc}
  \toprule
 & $\text{reg}, \text{imp}$ 
 & $\text{obj}, \text{imp}$ 
 & $\text{reg}, \text{prc}$ 
 & $\text{obj}, \text{prc}$ \\ 
  \midrule
  $\text{reg}, \text{imp}$ & 6.663 & 6.662 & 6.663 & 6.663 \\ 
  $\text{obj}, \text{imp}$ & 6.486 & 6.484 & 6.486 & 6.486 \\ 
  $\text{reg}, \text{prc}$ & 6.663 & 6.662 & 6.663 & 6.663 \\  
  $\text{obj}, \text{prc}$ & 6.663 & 6.662 & 6.663 & 6.663 \\ 
   \bottomrule
\end{tabular}                    
\caption{Average evolutionary fitness from Monte Carlo simulations of 100,000 symmetric $2 \times 2$ games}
\label{tab:ExpectedFitness_4Types}
\end{table}

Simulation results obviously reflect Fact~\ref{fact:equivalence2x2} in that all encounters in
which types $(\text{reg}, \text{imp})$, $(\text{reg}, \text{prc})$ or
$(\text{obj}, \text{prc})$ are substituted for one another yield identical results. More
interestingly, Table~\ref{tab:ExpectedFitness_4Types} shows that the maxmin strategy
$(\text{obj}, \text{imp})$ is strictly dominated by the three other strategies: in each column
(i.e., for each kind of opponent), the maxmin strategy is strictly worse than any of the other
three competitors. This has a number of interesting consequences.

Let's begin with a restricted scenario and look at more and more complex cases afterwards. If
we restrict attention to only imprecise belief types, then a state in which every agent has
regret-based preferences is the only \emph{evolutionarily stable state}. More strongly, since
$(\text{obj}, \text{imp})$ is strictly dominated by $(\text{reg}, \text{imp})$, we expect
selection that is driven by expected fitness to invariably weed out $(\text{obj}, \text{imp})$,
the maxmin choice rule, in favor of $(\text{reg}, \text{imp})$, regret minimization. In other
words, when taking average payoffs over the class of random games considered here, regret
minimization is a better classic choice rule than the maxmin rule from an evolutionary point of
view, despite the fact that the former, but not the latter, entertains non-veridical subjective
preferences.

Next, if we look at the competition between all four types represented in
Table~\ref{tab:ExpectedFitness_4Types}, $(\text{reg}, \text{imp})$ is no longer evolutionarily
stable. Rather, given behavioral equivalence (Fact~\ref{fact:equivalence2x2}), types
$(\text{reg}, \text{imp})$, $(\text{reg}, \text{prc})$, and $(\text{obj}, \text{prc})$ are all
\emph{neutrally stable} \citep{Maynard-Smith1982:Evolution-and-t}. But since
$(\text{obj}, \text{imp})$ is strictly dominated and so disfavored by fitness-based selection,
we are still drawn to conclude that maxmin behavior is weeded out in favor of a population with
a random distribution of the remaining three types.

Simulation results of the (discrete time) \emph{replicator dynamics}
\citep{TaylorJonker1978:Evolutionary-St} indeed show that random initial population
configurations are attracted to states with only three player types:
$(\text{reg}, \text{imp})$, $(\text{reg}, \text{prc})$ and $(\text{obj}, \text{prc})$. The
relative proportions of these depend on the initial population. This variability disappears if
we add a small mutation rate to the dynamics. Take a fixed, small mutation rate $\epsilon$ for
the probability that a player's preference type \emph{or} her belief type changes to another
preference type or belief type. The probability that a player randomly mutates into a
completely different player with altogether different preference type and belief type would
then be $\epsilon^2$. With these assumptions about ``component-wise mutations'', numerical
simulations of the (discrete time) \emph{replicator mutator dynamics}
\citep{Nowak2006:Evolutionary-Dy} show that already for very small mutation rates almost all
initial populations converge to a single fixed point in which the majority of players are
regret types. For instance, with $\epsilon = 0.001$, almost all initial populations are
attracted to a final distribution with proportions:\todo{recalculate these numbers for the
  actual 4x4 game!}

\begin{center}
  \begin{tabular}{cccc}
    $(\text{reg}, \text{imp})$ & $(\text{obj}, \text{imp})$ & $(\text{reg},
      \text{prc})$ & $(\text{obj}, \text{prc})$ \\ \hline
    0.289  & 0.021 &   0.398 &    0.289 
  \end{tabular}
\end{center}

What this suggests is that, if biological evolution selects behavior-generating mechanisms, not
behavior as such, it need not be the case that behaviorally equivalent mechanisms are treated
equally all the while. If mutation probabilities are a function of individual components (i.e.,
preference and belief types) of such behavior-generating mechanisms, it can be the case that
certain components are more strongly favored by a process of random mutation and
selection. This is exactly the case with regret-based subjective preferences in the present
example. Since regret-based subjective preferences are much better in connection with imprecise
beliefs than veridical preferences are, the proportion of expected regret minimizers,
$(\text{reg}, \text{prc})$, in the attracting state is substantially higher than that of
expected utility maximizers, $(\text{obj}, \text{prc})$, even though these types are
behaviorally equivalent.

@@

Results based on the single meta-game in Table~\ref{tab:ExpectedFitness_4Types}, which was
obtained by Monte Carlo simulation, are not fully general and possibly spoiled by random
fluctuations in the sampling procedure. Fortunately, for the case of $2 \times 2$ symmetric
games, the main result that maxmin types $(\text{obj}, \text{imp})$ are strictly dominated can
also be shown analytically for generous general conditions on how likely randomly sampled games
are.

\begin{proposition} \label{proposition1}
  Let $\mathcal{G}$ be the class of $2 \times 2$ symmetric games with payoffs sampled from a
  set of i.i.d.~values with at least three elements in the support. Then,
  $(\text{reg}, \text{imp})$ strictly dominates $(\text{reg}, \text{prc})$ in the resulting
  meta-game.
\end{proposition}

\begin{corollary} \label{corollary1} Considering only $(\text{obj}, \text{imp})$ and
  $(\text{reg}, \text{imp})$, and letting $\mathcal{G}$ be the class of $2 \times 2$ symmetric
  games with payoffs sampled from a set of i.i.d.~values with at least three elements in the
  support, the unique evolutionarily stable state is a monomorphic population of
  $(\text{reg}, \text{imp})$-players.
\end{corollary}
